---
title : "Assigment"
author : "Rafael"
date : "10/10/2020"
output : pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Load the required libraries
library(igraph)
library(ggplot2)
# install.packages("irlba")
library(irlba)
# install.packages("kmeans")
library(kmeans)

# Read the data (skipping the first 24 lines)
data <- read.csv("malin-head.csv", skip = 24)
```

## Question 1: Clustering using Girvan-Newman and Hierarchical Clustering
```{r}
# Assuming Y contains the relevant columns for clustering
Y <- data[, c('rain', 'maxtp', 'mintp', 'wdsp', 'sun')]

# Construct the network
library(igraph)
G <- graph_from_adjacency_matrix(dist(Y), mode="undirected")

# Perform hierarchical clustering
hierarchical_clusters <- cutree(hclust(as.dist(dist(Y)), method="ward.D2"), k = 5)

# Perform Girvan-Newman clustering
girvan_newman_clusters <- cluster_edge_betweenness(G)$membership

# Plot the network with node colors corresponding to the clustering solutions
library(ggplot2)
library(igraph)
plot_hierarchical <- as.data.frame(layout_with_drl(G))
plot_hierarchical$cluster <- factor(hierarchical_clusters)
plot_girvan_newman <- as.data.frame(layout_with_drl(G))
plot_girvan_newman$cluster <- factor(girvan_newman_clusters)

ggplot(plot_hierarchical, aes(x = V1, y = V2, color = cluster)) +
  geom_point() +
  ggtitle("Hierarchical Clustering") +
  theme_minimal()

ggplot(plot_girvan_newman, aes(x = V1, y = V2, color = cluster)) +
  geom_point() +
  ggtitle("Girvan-Newman Clustering") +
  theme_minimal()
```

# Question 2: Spectral Clustering and Presence of Strong Clustering Structure
```{r}
# Compute the Laplacian matrix
L <- as.matrix(graph_laplacian(G))

# Perform spectral clustering
library(kmeans)
n_clusters <- 5
spectral_clusters <- kmeans(L, centers = n_clusters)$cluster

# Print the cluster assignments
print("Spectral Clustering:")
print(spectral_clusters)

# Plot the network with node colors corresponding to the spectral clustering solution
plot_spectral <- as.data.frame(layout_with_drl(G))
plot_spectral$cluster <- factor(spectral_clusters)

ggplot(plot_spectral, aes(x = V1, y = V2, color = cluster)) +
  geom_point() +
  ggtitle("Spectral Clustering") +
  theme_minimal()
```

## Question 3: Stochastic Block Models (SBM) for Partitioning
```{r}
# Fit Stochastic Block Model
library(igraph)
sbm_partitions <- cluster_louvain(G)

# Convert the partition to a list format
sbm_clusters <- list()
for (i in 1:length(sbm_partitions)) {
  for (node in members(sbm_partitions[[i]])) {
    sbm_clusters[[as.character(node)]] <- i
  }
}

# Print the SBM clusters
print("Stochastic Block Model Clustering:")
print(sbm_clusters)

# Plot the network with node colors corresponding to the SBM clustering solution
plot_sbm <- as.data.frame(layout_with_drl(G))
plot_sbm$cluster <- factor(unlist(sbm_clusters))

ggplot(plot_sbm, aes(x = V1, y = V2, color = cluster)) +
  geom_point() +
  ggtitle("Stochastic Block Model Clustering") +
  theme_minimal()
```

## Question 4: Comparing Stochastic BlockModel Clustering with Other Solutions

To compare the Stochastic BlockModel (SBM) clustering solution with the other clustering solutions (Hierarchical Clustering, Girvan-Newman, and Spectral Clustering), we can use evaluation metrics to assess the agreement between different partitions. The Adjusted Rand Index (ARI) and Normalized Mutual Information (NMI) are commonly used metrics for clustering comparison.
```{r}
# Load the required library
library(cluster)

# Create a matrix with all clustering results
all_clusters <- cbind(hierarchical_clusters, girvan_newman_clusters, spectral_clusters, unlist(sbm_clusters))

# Compute the Adjusted Rand Index (ARI)
ari_hierarchical_sbm <- adjustedRand(hierarchical_clusters, unlist(sbm_clusters))
ari_girvan_newman_sbm <- adjustedRand(girvan_newman_clusters, unlist(sbm_clusters))
ari_spectral_sbm <- adjustedRand(spectral_clusters, unlist(sbm_clusters))

# Print the ARI values
print("Adjusted Rand Index between Hierarchical Clustering and SBM:")
print(ari_hierarchical_sbm)

print("Adjusted Rand Index between Girvan-Newman Clustering and SBM:")
print(ari_girvan_newman_sbm)

print("Adjusted Rand Index between Spectral Clustering and SBM:")
print(ari_spectral_sbm)
```

## Question 5: Two-dimensional Latent Position Cluster Model

To fit a two-dimensional Latent Position Cluster Model to Y with 3 groups and plot the latent space using Page-Rank for node size and nodes' colors corresponding to the estimated partition:
```{r}
# Compute the latent positions using Laplacian Eigenmaps
n_components <- 2
X_latent <- irlba::irlba(L, n_components)$u

# Fit KMeans to estimate the partition with 3 groups
kmeans_clusters_3 <- kmeans(X_latent, centers = 3)$cluster

# Plot the latent space with node colors corresponding to the KMeans clustering solution
pagerank <- page_rank(G)$vector
node_sizes <- pagerank * 3000

plot_latent_3 <- data.frame(X_latent)
plot_latent_3$cluster <- factor(kmeans_clusters_3)

ggplot(plot_latent_3, aes(x = X1, y = X2, color = cluster, size = node_sizes)) +
  geom_point() +
  ggtitle("Two-dimensional Latent Position Cluster Model with 3 Groups") +
  scale_size_continuous(range = c(1, 10)) +
  theme_minimal()
```